---
title: "STAT4600: Assignment 1"
author: "Xinhao Wang"
date: 'Due: Tuesday, January 16, 2018'
output: html_document
---
As was discussed in class, the goal of this first assignment is to help you become more familiar with R Markdown.

Specifically, you have to reproduce this document in its entirety, including  
- the title (but use your name as the author),  
- the instructions,  
- the text (properly *italicized*, **bolded**, etc.),  
- R code,  
- R output, including graphs,  
- mathematical expressions and formulas.  

In other words, you are asked to prepare an RMarkdown file that reproduces this document exactly, except for:  
- having your name at the top as the author,  
- the computation times in the table at the end of this document will be different depending on your system,  
- your personnal answers to the two questions appearing at the end of the document.  

Your RMarkdown file should be clean and as simple as possible.  

## Matrix multiplication in R: the basic idea
We know that when two matrices $A:m \times k$ and $B:k \times n$ are multiplied, the result is another matrix C:m×n with elements given by
$$
c_{ij} = \sum_{k=1}^n{a_{ik}b_{kj}}.
$$
For this, R has the predefined function `%*%`. In order to experiment with it, we define the following two matrices.
```{r} 
A <- matrix(c(2,5,6,1,1,0), nrow = 3, ncol = 2, byrow = T)
A
B <- matrix(c(0,0,4,6,2,3,1,1,1,5), nrow = 2, ncol = 5, byrow = T)
B
```
The product of `A` and `B`, which is another matrix of dimensions 3×5, is then obtained as
```{r}
A %*% B             # We will call this method 1
```
We note that the product `B %*% A` cannot be calculated because of non-matching dimensions.

This product can also be obtained using a full blown-out brute-force approach, which would be very natural and efficient using many other programming languages, but that should be avoided in R:

```{r}
C <- matrix(0, nrow = 3, ncol = 5)            # We will call this method 2
for(i in 1:3){
  for(j in 1:5){
    for(k in 1:2){
      C[i,j] <- C[i,j] + A[i,k]*B[k,j]
    }
  }
}
C
```

A slightly better approach, as it uses only two loops:
```{r}
C <- matrix(0, nrow = 3, ncol = 5)            # We will call this method 3
for(i in 1:3){
  for(j in 1:5){
      C[i,j] <- sum(A[i,]*B[,j])
  }
}
C
```

It is also possible to use only one loop (by using the underlying vector associated to `C`):
```{r}
C <- matrix(0, nrow = 3, ncol = 5)            # We will call this method 4
for(k in 1:15){
  j <- ceiling(k/3)                           # this is the column number
  i <- k - (j-1)*3                            # this is the row number
  C[k] <- sum(A[i,]*B[,j])
}
C
```

But, none of these other approaches are nearly as efficient as the prefined function `%*%`.

###A simple but convincing experiment
In order to see the difference in efficiency between these methods, we define new very large matrices `A` and `B` (with elements generated at random) and try the four methods to calculate their product. The code presented above will need to be adapted for this.
```{r,eval=FALSE}
A <- matrix(sample.int(n=10,size=45000,replace=T), nrow = 300, ncol = 150)
B <- matrix(sample.int(n=10,size=15000,replace=T), nrow = 150, ncol = 100)
```

Then, using the `system.time` function and running the four methods above to obtain the product of `A` and `B`, we can make the following table.

: Computation times for the different methods

```{r,echo=FALSE}

A <- matrix(sample.int(n=10,size=45000,replace=T), nrow = 300, ncol = 150)
B <- matrix(sample.int(n=10,size=15000,replace=T), nrow = 150, ncol = 100)

method1 <- system.time(A %*% B)
time.m1 <- method1["elapsed"]



A <- matrix(sample.int(n=10,size=45000,replace=T), nrow = 300, ncol = 150)
B <- matrix(sample.int(n=10,size=15000,replace=T), nrow = 150, ncol = 100)

method2 <- system.time( 
  {C <- matrix(0, nrow = 300, ncol = 100)            # We will call this method 2
for(i in 1:300){
  for(j in 1:100){
    for(k in 1:150){
      C[i,j] <- C[i,j] + A[i,k]*B[k,j]
    }}}}
)
time.m2 <- method2["elapsed"]



A <- matrix(sample.int(n=10,size=45000,replace=T), nrow = 300, ncol = 150)
B <- matrix(sample.int(n=10,size=15000,replace=T), nrow = 150, ncol = 100)

method3 <- system.time(
  {C <- matrix(0, nrow = 300, ncol = 100)            # We will call this method 3
for(i in 1:300){
  for(j in 1:100){
      C[i,j] <- sum(A[i,]*B[,j])
  }}}
)
time.m3 <- method3["elapsed"]



A <- matrix(sample.int(n=10,size=45000,replace=T), nrow = 300, ncol = 150)
B <- matrix(sample.int(n=10,size=15000,replace=T), nrow = 150, ncol = 100)

method4 <- system.time(
  {
    C <- matrix(0, nrow = 300, ncol = 100)            # We will call this method 4
for(k in 1:30000){
  j <- ceiling(k/300)                           # this is the column number
  i <- k - (j-1)*300                            # this is the row number
  C[k] <- sum(A[i,]*B[,j])
}
  }
)
time.m4 <- method4["elapsed"]
```


|Method|Elapsed time (sec)|
|:----------------:|:--------:|
|1      |    `r time.m1` |
|2      |    `r time.m2` |
|3      |    `r time.m3` |
|4      |    `r time.m4` |

###Two Short Questions

*Question 1.*

Explain what the code `matrix(sample.int(n=10,size=45000,replace=T),nrow = 300, ncol = 150)` does.

`Answer`: creating a 300*150 matrix,so there is 45000 elements in this matrix. and for each element random from 1 to 10           with repalcement.

*Question 2.*

`Briefly` explain (at most a few lines for each) how method 2,3 and 4 work.

`Answer`:  

`method 2`: 
First we make a $3\times5$ matrix names 'C' with each element is 0, and we use 3 'for' loop to calculate matrix A and matrix B. For every numbers in A and B, we calculate each number one by one,and add them together to get an element in C. For example, Using first number in A's first row $\times$ first number in B's first column, then second number in A's first row $\times$ second number in B's first column and so on. And using C[i,j] <- C[i,j] + A[i,k]*B[k,j] to add those numbers together until we get all those 15 numbers in C.

`method 3`:
It is like we use every rows in A times very columns in B to get each number in C. For example, we using A[1,]$\times$B[,1] which is first row in A times first col in B, Then the answer is a vector with some numbers. and using sum(A[i,]$\times$B[,j]) to add all those number to get $C_{11}$. same thing for $C_{12}$ to $C_{35}$.

`method 4`:
for this method we using 'ceiling' to rounding of numbers to get colcumn numbers, and using formula 'i <- k - (j-1)*3' to define row numbers. once we got the row numbers and column numbers, then the calculate step is same with method 3.